WDIAS : A Microservices-Based Weather Data Integration and Assimilation System\\

### 2020-04-26 glossary.tex
Removed commercial database names
Problem statement
Moved k8s to architectural decisions
Updated -> Method - Microservices

\newacronym{fews}{Delft-FEWS} {Delft-FEWS, Deltares}
\newacronym{lead}{LEAD}{Linked Environments for Atmospheric Discovery}
\newacronym{dias}{DIAS}{Data Integration and Assimilation System}
\newacronym{madis}{MADIS}{Meteorological Assimilation Data Ingest System}

\newacronym{nwm}{NWMs}{Numerical Weather Models}
\newacronym{netCDF}{netCDF}{Network Common Data Form}
\newacronym{soa}{SOA}{Service Oriented Architecture}
\newacronym{wrf}{WRF}{Weather Research and Forecast}
\newacronym{esb}{ESB}{Enterprise Service Bus}
\newacronym{microservice}{Microservice}{Microservice Architecture}

\newacronym{mysql}{MySQL}{MySQL}
\newacronym{influxdb}{InfluxDB}{InfluxDB}
\newacronym{redis}{Redis}{Redis}
\newacronym{mongodb}{MongoDB}{MongoDB}
\newacronym{k8s}{K8s}{Kubernetes}
\newacronym{eks}{AWS EKS}{Amazon Elastic Kubernetes Service}

\newacronym{go}{Go Lang}{Go Programming Language}
\newacronym{rps}{RPS}{Requests Per Second}
\newacronym{api}{API}{Application Programming Interface}
\newacronym{jmeter}{JMeter}{Apache JMeter}
\newacronym{rps}{RPS}{Request per Second}

\newacronym{curw}{CUrW SL}{Urban Center for Water, Sri Lanka}

>>>>

\newacronym{fews}{Delft-FEWS} {Deltares FEWS}
\newacronym{lead}{LEAD}{Linked Environments for Atmospheric Discovery}
\newacronym{dias}{DIAS}{Data Integration and Assimilation System}
\newacronym{madis}{MADIS}{Meteorological Assimilation Data Ingest System}

\newacronym{nwm}{NWMs}{Numerical Weather Models}
\newacronym{soa}{SOA}{Service Oriented Architecture}
\newacronym{wrf}{WRF}{Weather Research and Forecast}
\newacronym{esb}{ESB}{Enterprise Service Bus}
\newacronym{microservice}{Microservice}{Microservice Architecture}

\newacronym{netCDF}{netCDF}{Network Common Data Form}
\newacronym{GRIB}{GRIB}{General Regularly-distributed Information in Binary form}
\newacronym{csv}{CSV}{Comma-separated Values}

\newacronym{rdbms}{RDBMS}{Relational Database Management System}
\newacronym{k8s}{K8s}{Kubernetes}
\newacronym{eks}{Amazon EKS}{Amazon Elastic Kubernetes Service}

\newacronym{go}{Go Lang}{Go Programming Language}
\newacronym{rps}{RPS}{Requests Per Second}
\newacronym{api}{API}{Application Programming Interface}
\newacronym{rest}{REST}{Representational State Transfer}

\newacronym{curw}{CUrW-SL}{Center for Urban Water, Sri Lanka}

### 2020-04-26 method/sec_0_introduction.tex
Updated -> Method - Microservices

This chapter presents the design decisions taken while implementing \acrfull{wdias}, and how is the system got \hl{adapted from the systems analyzed in Chapter} \ref{ch:literature}.
\cref{se:high_level_design} presents a brief discussion on how the weather forecasting is happening and why it needs to interact with multiple data formats.
\cref{se:architectural_decisions} presents how the architectural design of proposed solution evolved while designing and implementing \acrshort{wdias}.
\cref{se:microservice} presents %gives a better understanding of the 
microservice architectural concepts followed while implementing \acrshort{wdias}.
Then the reasons behind developing \acrshort{wdias} database structure and technologies used are discussed in \cref{se:db_struct}.
\cref{se:data_preprocess} discusses how \acrshort{wdias} enables data preprocessing and adding capabilities using extensions.
Finally, \cref{se:query} includes details on how to perform timeseries search and geo-based queries on the \acrshort{wdias}.

\dbc{It's just WDIAS. "the" should be used only on essential places to being the meaning. }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{High-level Design}
\label{se:high_level_design}

To get an overall idea of the flow of weather prediction, let us look into one of the \hl{flows using} at the \acrshort{curw} \cite{CUrWSL2017SL}.

\begin{enumerate}
    \item Extract \acrshort{GRIB} \acrshort{wrf} model data to \acrshort{netCDF}
    \item \hl{From} \acrshort{netCDF} to \acrshort{csv} of each location data, then feed to HEC-HMS hydro-dynamic model and get the output as a \acrshort{csv} file.
    \item \acrshort{csv} and \hl{Rain cell Grid feed} into the FLO2D hydrologic model, and get the water level as ASCII Grid files after the simulation is complete.
    \item Upload ASCII files to \hl{ArcGIS create raster files, then use for the LOSS Estimation}
\end{enumerate}

\hl{When looked into the above flow, it gives a clear understanding of multiple data formats using while weather predictions.}
Also, to correct the model outputs, \hl{it required} to collect real-time data such as:
\begin{itemize}
    \item Frequently collected automated weather station data via HTTP protocol
    \item Frequently collected automated water level stations data
\end{itemize}

\hl{Other than above many data formats using for handling data based on devices and standards which are using.} Also, for data modeling, it needs to convert the data \hl{into model compatible data format}. Then \hl{after successfully run the model it needs to collect data from the model output data formats.}

@@

\subsubsection{Integration}
The system should capable of integrating data from different sources such as satellite data, high end, and low-end weather station, etc. And the system should be able to handle multidimensional spatial and temporal weather data efficiently and optimally. 
\subsubsection{Assimilation}
Then the system should be able to fulfill weather models varying data requirements. Also, those models reproduce a large set of redundant data, thus the system should store the bulk data while optimizing the disk space.
\subsubsection{Dissemination}
Then different users should be able to retrieve data as they required. Also, the users should be able to easy to search into the available that in the system base on timeseries metadata or based on Geo queries.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \paragraph{Timeseries}-- A timeseries is simply a series of data points ordered in time. In weather domain, it interest in timeseries in perspective of observations to forecasting. Each timeseries, the data points can be form in different formats as well. For example, scalar (0D), vector (1D), grid (2D), and polygon (2D).

% In the proof of concept design described below, WDIAS focuses on handling scalar, vector and grid timeseries data. However, the system could be extended to handle polygon-based timeseries as well.

% \subsubsection{Base of WDIAS architecture}
% \hl{The base of WDIAS architecture origins based on attributes of a timeseries.} There are many attributes to differentiate timeseries one from another. But 
% Among many attribute to differentiate a timeseries from another, the following can be consider as the key attributes in uniquely identify a timeseries from another: % for the \acrshort{wdias}.

% \begin{itemize}
%     \item Module -- module that generates the timeseries
%     \item Value type -- Scalar, vector, grid, or polygon
%     \item Location of readings
%     \item Parameter -- measuared parameter
%     \item hl{Timeseries type}
%     \begin{itemize}
%         \item Sources -- External or simulated
%         \item Category -- Historical or forecast
%     \end{itemize}
%     \item Time step
% \end{itemize}
% \dbc{Use bullets only when essential. Other places use paragraphs.}

% A detail description of the above attributes are presented in \cref{se:db_struct}.

>>>>

This Chapter presents the design decisions taken while implementing \acrfull{wdias} and how we adopt from the systems analyzed in \ref{ch:literature}.
\cref{se:high_level_design} presents a brief discussion on how  the weather forecasting is happening and why does it need to interact with multiple data formats.
\cref{se:architectural_decisions} shows how the architectural design of the proposed solution evolved while designing and implementing \acrshort{wdias}.
\cref{se:microservice} presents microservice architectural concepts followed while developing the \acrshort{wdias}. Then the reasons behind developing \acrshort{wdias} database structure and technologies used are discussed in  \cref{se:db_struct}.
\cref{se:data_preprocess} discusses how \acrshort{wdias} enables data preprocessing and adding capabilities using extensions. Finally, \cref{se:query} includes details on how to perform timeseries search and geo-based queries on the \acrshort{wdias}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{High-level Design}
\label{se:high_level_design}

To get an overall idea of the flow of weather prediction, let us look into one of the weather forecastings flows using the \acrshort{curw} \cite{CUrWSL2017SL}.

\begin{enumerate}
    \item Extract \acrshort{GRIB} \acrshort{wrf} model data to the \acrshort{netCDF} format
    \item Extract \acrshort{netCDF} data to \acrshort{csv} format for each location data, then feed to HEC-HMS hydrodynamic model and get the output as a \acrshort{csv} file.
    \item HEC-HMS's \acrshort{csv} output and precipitation as Grid files feed into the FLO2D hydrologic model and get the water level as ASCII Grid files after the simulation is complete.
    \item Upload above ASCII files to a geographic information system after convert to a compatible format, then use for the loss estimation
\end{enumerate}

In the above weather forecasting steps, we can see that the weather data need to convert to different formats while feeding from one model or step to another. After the models run successfully, the resulting output data may not be accurate or biased based on different factors. In such situations, we need to correct predicted weather data by comparing them with accurate real-time weather data. So, we are required to collect real-time data such as:
\begin{itemize}
    \item Frequently collect automated weather station data via HTTP protocol
    \item Collect accumulated weather data over a given period by automatic water level stations
\end{itemize}

Most weather data collecting devices use their data formats and different standard protocols to transmit the data. Then we need to validate the data and modify the data to remove errors. Also, these data need to convert into model compatible data format before feeding into models. The models output data in different formats and need to store those bulk data efficiently, and the system should be flexible enough to upgrade to support new data formats easily. Essentially, the system should share the real data, forecasted data, and other created data with relevant parties based on authorization.

@@

\subsubsection{Integration }
The system should be capable of integrating data from different sources such as satellite data, high end, and low-end weather station. And the system should be able to handle multidimensional spatial and temporal weather data efficiently and optimally.
\subsubsection{Assimilation}
Then the system should be able to fulfill weather models varying data requirements. Also, those models reproduce a large set of redundant data. Thus the system should store the bulk data while optimizing the disk space.
\subsubsection{Dissemination}
Then different users should be able to retrieve data as they require. Also, the users should be able to easily search into the available information in the system based on timeseries metadata or based on Geo queries.

### 2020-04-26 method/sec_1-architectural_decisions.tex
Updated -> Method - Microservices

The initial design of \acrshort{wdias} used the \acrfull{soa}. Moreover, it tried to use an \acrfull{esb} to integrate different modules, as \acrshort{esb} could act as a \hl{common layer} for all of the modules. By default, \acrshort{esb} also provides publisher/subscribe (pub/sub) capabilities.
\hl{Because \acrshort{esb} mediators can be used to integrate the modules which are described in Section} \ref{subse:modules_weather_data_integration_sys}. 
\dbc{Last sentence incomplete too.}

The \acrshort{esb} has the processing units called \emph{mediators}. A mediator can take a message, carry out some predefined actions on it, and output the modified message. These mediator capabilities can use to integrate the modules of a weather data system as shown in \cref{fi:wdia_components}. \acrshort{esb} also supports different transportation protocols such as HTTP and Web sockets. However, it turned out that \acrshort{esb} is not suitable for data streaming or bulk data processing. Also, \acrshort{esb} suffers from a single point of failure, as all the messages are going through a common bus and slower communication (cant use for transfer data).

In the second design phase of \acrshort{wdias}, we attempted to use the actor model using the AKKA framework \cite{HewittWhyModel} to overcome above \acrshort{esb} drawbacks. At the same time, we moved away from the \acrshort{soa} architecture design to the microservice architecture design. Thus we redesign the system architecture with decomposing into microservices and used actors to implement each microservice in the system. In that case, an actor or an actor with slave actors was a map to each microservice in \acrshort{wdias}. \acrshort{soa} focuses on imperative programming, while the microservices architecture focuses on a reactive actor programming style using faster messaging mechanisms.

When compare to \acrshort{soa}, the microservice architecture has several advantages such as the following:
\begin{itemize}
    \item Follow the principle of a single responsibility.
    \item Resilient / flexible because a failure of a service has no influence on other services. If you have monolithic or major service errors in a service / module, this can have an impact on other modules / functionalities.
    \item High scalability because demanding services can be implemented on different servers to improve performance and stay away from other services so that they do not affect other services. It will be difficult to achieve the same with a single large monolithic service.
    \item Easy to improve because the dependencies are less. It is also easier to change and test.
    \item Little impact on other services because each service is independent.
    \item Easy to understand because they represent a small functionality.
    \item Ease of deployment.
\end{itemize}

% /hl{The typical \acrshort{soa} model, for example, usually has more dependent \acrshort{esb}s, with microservices using faster messaging mechanisms. \acrshort{soa} also focuses on imperative programming, whereas microservices architecture focuses on a responsive-actor programming style.}
\dbc{Multiple design rounds and decisions are mixed in the para. First finish discussing ESB. Then finish actor model in another para. Then focus on mircoservice details.}

Even with using the AKKA framework, it has some of the disadvantages of implementing each microservice as an actor as described in AKKA documentation \cite{Akka.ioWhenCluster}. One of the attractive features of microservice architecture is the independent nature of the microservices. This allows choosing a different technology for each microservice based on the advantage that could bring in. Also allowed to independently develop and maintain by multiple smaller and more focused teams/communities. But using the actors as microservice, the actors messaging between different services cause to result in a too-tight code coupling between the services and difficulties deploying these independent of each other, which is one of the main reasons for using a microservices architecture \cite{Akka.ioWhenCluster}. To overcome these issues, we moved to the concept of container-orchestration based architecture.

% \begin{itemize}
%     \item \hl{Unable to choose technology. But} \acrshort{k8s} \hl{allows you to choose technology that is best suited for a particular functionality}
%     \item \hl{As mentioned in the AKKA documentation} \cite{Akka.ioWhenCluster}, \hl{it is better for developing one microservice within the whole} \acrshort{wdias} design.
% \end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsection{Brief introduction to the \acrfull{k8s}}
% \label{sebse:k8s_intro}

Containerization is an alternative to virtualization that supports virtual machines or hypervisors. This includes encapsulating or grouping the software code and all its dependencies so that it can work uniformly and consistently in any infrastructure. \cite{IBMContainerizationExplained}. The concept of containerization and process isolation has emerged after the open-source Docker Engine \cite{DockerAppContainerization}, an industry standard for package software into standardized units for development, shipment, and deployment.

The \acrshort{wdias} is using \acrfull{k8s} as the container orchestration system which is an open-source tool for automating deployment, scaling, and management of containerized applications \cite{LinuxFoundationProduction-GradeKubernetes}. It groups the containers that make up an application into logical units for easy management and detection. While implementing the \acrshort{wdias}, we used \acrshort{k8s} as the container-orchestration system, to deploy each microservice as container and manage with providing scalability and availability.

\begin{figure}[htp]
    \centering
    \includegraphics[width=1\textwidth]{method/microservice/k8s_architecture_v3.jpg}
    \caption{\acrfull{k8s} architecture.}
    \label{fi:k8s_architecture}
\end{figure}
\cref{fi:k8s_architecture} gives an overall idea on the components of \acrshort{k8s},
\begin{itemize}
    \item pods -- Cluster of containers that can group other container images in a single unit.
    \item nodes -- The machines (VMs, physical servers, etc) in a cluster that runs your applications and cloud workflows.
    \item kubelet -- An agent that runs on each node in the cluster. It makes sure that containers are running in a pod.
    \item kube-proxy -- A network proxy that runs on each node in the cluster, and maintains network rules on nodes.
    \item kubernetes master -- Responsible for maintaining the desired state for your cluster.
    \item etcd -- Consistent and highly-available key value store used as Kubernetes’ backing store for all cluster data.
\end{itemize}

Users can add much as required Nodes into the \acrshort{k8s} cluster, and \acrshort{k8s} manage and deploy applications as pods into cluster nodes.
Each microservice can deploy as a pod which is a group of containerized applications. And \acrshort{k8s} able to scale as much as the user wants as described above.
By separating each microservice as a container, and running them as multiple pods inside the \acrshort{k8s} removes the tight coupling of microservices when compared to the actors. This preserves the independent nature of the microservices which allows to high scale the system.

>>>>

The initial design of \acrshort{wdias} used the \acrfull{soa}. Moreover, it tried to use an \acrfull{esb} to integrate different modules, as \acrshort{esb} could act as a shared layer for all of the modules to interact together. By default, \acrshort{esb} also provides publisher/subscribe (pub/sub) capabilities.

The \acrshort{esb} has the processing units called \emph{mediators}. A mediator can take a message, carry out some predefined actions on it, and output the modified message. These mediator capabilities can be used to integrate the modules of a weather data system, as shown in \cref{fi:wdia_components}. \acrshort{esb} also supports different transportation protocols such as HTTP and Web sockets. However, we realized that \acrshort{esb} is not suitable for data streaming or bulk data processing. Also, \acrshort{esb} suffers from a single point of failure, as all the messages are going through a common shared bus and slower communication (can not be used for transfer data). In the second design phase of \acrshort{wdias}, we attempted to use the actor model using the AKKA framework \cite{HewittWhyModel} to overcome above \acrshort{esb} drawbacks. At the same time, we moved away from the \acrshort{soa} architecture design to the microservice architecture design.

Thus we redesigned the system architecture with decomposing into microservices and used actors to implement each microservice in the system. In that case, we mapped each microservice in \acrshort{wdias} to an actor or an actor with slave actors. \acrshort{soa} focuses on imperative programming, while the microservices architecture focuses on a reactive actor programming style using faster messaging mechanisms.
When compared to \acrshort{soa}, the microservice architecture has several advantages:
\begin{itemize}
    \item Follow the principle of a single responsibility.
    \item Resilient/flexible because a failure of service does not influence other services. If you have monolithic or significant service errors in a service/module, this can have an impact on other modules/functionalities.
    \item High scalability because demanding services can be deployed on different servers to improve performance and stay away from other services so that they do not affect other services. It will be challenging to achieve the same with a single large monolithic service.
    \item Easy to improve the system because the microservices are decoupled from other services. It is also easier to change and test each microservices.
    \item Little impact on other services because each service is independent.
    \item Easy to understand because they represent a small functionality.
    \item Ease of deployment.
\end{itemize}

The AKKA framework has some of the disadvantages of implementing each microservice as an actor, as described in AKKA documentation \cite{Akka.ioWhenCluster}. One of the attractive features of microservice architecture is the independent nature of the microservices. The above concept allows choosing different technologies for each microservice based on the advantage that could bring in. Also, the microservice architecture will enable us to independently develop and maintain a system by multiple smaller and more focused teams/communities. But using the actors as microservice, the actors communicate using message passing cause to result in a too-tight code coupling between the services and difficulty to deploy services independent of each other, which is one of the main reasons for using a microservices architecture \cite{Akka.ioWhenCluster}. To overcome these problems, we moved to the concept of container orchestration based microservice architecture.

Containerization is an alternative to virtualization that supports virtual machines or hypervisors. It includes encapsulating or grouping the software code and all its dependencies so that it can work uniformly and consistently in any infrastructure \cite{IBMContainerizationExplained}. The concept of containerization and process isolation has emerged after the open source Docker Engine \cite{DockerAppContainerization}, an industry standard for package software into standardized units for development, shipment, and deployment.

The \acrshort{wdias} used \acrfull{k8s} as the container orchestration system. \acrshort{k8s} is an open-source tool for automating deployment, scaling, and management of containerized applications [15]. It groups the containers that make up an application into logical units for easy management and detection. Using \acrshort{k8s}, we deploy each microservice as a container and manage it by providing scalability and availability.

\begin{figure}[htp]
    \centering
    \includegraphics[width=1\textwidth]{method/microservice/k8s_architecture_v3.jpg}
    \caption{\acrfull{k8s} architecture.}
    \label{fi:k8s_architecture}
\end{figure}

\cref{fi:k8s_architecture} gives an overall idea on the components of \acrshort{k8s},
\begin{itemize}
    \item pods – Cluster of containers that can group other container images in a single unit.
    \item nodes – The machines (VMs, physical servers, etc) in a cluster that runs your applications and cloud workflows.
    \item kubelet – An agent that runs on each node in the cluster. It makes sure that containers are running in a pod.
    \item kube-proxy – A network proxy that runs on each node in the cluster, and maintains network rules on nodes.
    \item kubernetes master – Responsible for maintaining the desired state for your cluster.
    \item etcd – Consistent and highly-available key value store used as Kubernetes backing store for all cluster data.
\end{itemize}

Users can add much as Nodes required into the \acrshort{k8s} cluster, and \acrshort{k8s} manage and deploy applications as pods into cluster nodes. Each microservice can deploy as a pod, which is a group of containerized applications. And \acrshort{k8s} is able to scale as much as the user requirements by deploying multiple copies of the same pod. By separating each microservice as a container, and running them as multiple pods inside the \acrshort{k8s} removes the tight coupling of microservices when compared to the actors. It preserves the independent nature of the microservices and supports the high scalability of the system.
